---
title: "AIT Final Project"
author: "Sai Jaswanth Kumar Kunku"
date: "11/25/2021"
output:
  pdf_document: default
  html_document:
    df_print: paged
---
## Library Import
```{r,results='hide'}
library(tidyverse)
library(lubridate)
library(stringr)
library(reshape)
library(plotly)
library(psych)
library(corrplot)
library(ggcorrplot)
```

```{r}
spotify_songs <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-01-21/spotify_songs.csv')

write.csv(spotify_songs,"./spotifydata.csv")
```

#Data Import,cleaning and converting cleaned dataset into CSV for exploration in other platforms(Python,SQL)
```{r,eval=FALSE}
df <- read.csv('spotifydata.csv')
names(df)
#removing unwanted columns
df <- df[-c(1,5,6,8,9,11)]
names(df)
dim(df)
summary(df)
#Finding the Missing values
missingvalues=colSums(is.na(df))
missingvalues
#removing missing values
df<-na.omit(df)  #Checking duplicates
sum(duplicated(df)) #We have zero duplicates in the dataset
#extracting year from the date
df$track_album_release_date<-as.Date(df$track_album_release_date)
df$year<-format(as.Date(df$track_album_release_date), "%Y")
df$year<-as.numeric(df$year)
df$duration_ms<-df$duration_ms/(1000*60)
#names(df$duration_ms) <- 'duration(min)'
#replacing different versions of same song such as edit|remix|remastered|remaster|radio|version|original|mix to eliminate duplicates
df$track_name<-gsub("\\-.*|\\[.*|\\(.*","",df$track_name)
#Removing duplicates tracks to get final clean dataset
unique_rows <- !duplicated(df[c("track_name","track_artist")])
df <- df[unique_rows,]
df<-subset(df, year!="NA")
df<-as.data.frame(df)
#writing the new dataset
#write.table(df,"./final.csv",sep=";",row.names=FALSE,fileEncoding = "UTF-8")
song_features<-df %>% select(year,track_popularity,danceability,energy,key,loudness,mode,speechiness,instrumentalness,acousticness,liveness,valence,tempo)
```


#1. What is the average popularity of tracks that were released during that year?
```{r}
year_popularity<-df %>%group_by(year) %>% summarise_at(vars(track_popularity), list(avg_popularity = mean))
as.data.frame(year_popularity)
year_popularity%>%plot_ly(x = ~year, y = ~avg_popularity, type = 'scatter', mode = 'lines+markers') %>%
  layout(title='Average Popularity over the years',xaxis = list(title='Year'),yaxis=list(title='Popularity'))
```
#2. How the audio features such as key, mode etc. are distributed and correlated with each other?
```{r,fig.asp = 0.6,eval=FALSE}

cor(song_features)
#pairs.panels(song_features)

ggcorrplot(cor(song_features), hc.order = TRUE,show.diag = FALSE,lab=TRUE,insig = "blank",digits = 2,lab_size=3,colors= c("#6D9EC1", "white", "#E46726"))+ 
  labs(title = "Correlation between Song Features")
```
##Predicting popularity based on song_characterstics
```{r,eval=FALSE}
lm.mod = lm(track_popularity~instrumentalness+acousticness+liveness+tempo+energy+loudness+
              danceability+valence+year, data=df);
summary(lm.mod)

sample=sample(1:nrow(df),size=nrow(df)*.7)
train<-df[sample,] #Select the 80% of rows
test<-df[-sample,] #Select the 20% of rows
#Logistic regression Model on training data
model<-glm(lm.mod,data=train)
#Predicting the model on test data
predicted=predict(model,test)

actuals_preds <- data.frame(cbind(actuals=test$track_popularity, predicteds=predicted))

min_max_accuracy <- mean(apply(actuals_preds, 1, min) / apply(actuals_preds, 1, max)) 
min_max_accuracy*100
RMSE(predicted, test$track_popularity)

```

#3. How is the music industry is evolving over the period and mainly in the 21st Century?

```{r,eval=FALSE}
df %>% group_by(year)  %>% summarise(count = n()) %>% subset(year <= 2019) %>% plot_ly(x = ~year,y = ~count,type = 'scatter',mode='lines+markers')%>%
layout(title='Song Releases between 1957 and 2020',xaxis = list(title='Year'),yaxis=list(title='No of songs'))

```
```{r,eval=FALSE}
plot_zoom_year <- df %>% select(year) %>%group_by(year) %>%summarise(count = n()) %>% subset(year >= 2000)
plot_zoom_year%>%plot_ly(x = ~year,y = ~count,type = 'scatter',mode='lines+markers')%>%
  layout(title='Music in 21st Century',xaxis = list(title='Year'),yaxis=list(title='No of songs released'))

```


```{r,eval=FALSE}
y1=song_features%>%subset(year < 2000)
y2=song_features%>%subset(year >= 2000)
colMeans(y1[ , 1:ncol(y1)])
colMeans(y2[ , 1:ncol(y2)])
```

```{r,eval=FALSE}
library(ggplot2)
yeardata=df
yeardata$Music[df$year < 2000] <- "20th Century"
yeardata$Music[df$year > 2000] <- "21st Century"
yeardata<-na.omit(yeardata)
df3=yeardata %>% group_by(Music,playlist_genre) %>% summarise(count = n())
ggplot(df3, aes(y=count, x=playlist_genre)) + geom_bar(aes(fill = Music),position = "dodge", stat = "identity")+labs(
  title="Distribution of Songs in different Genres",
  x="Genre",
  y="Count"
)



```

